Distanced-based VOCC analysis

### 2019-05-17 analysis/VOCC/dist-vocc.Rmd
fix calcslope to always have time axis in single units

library(dplyr)
library(ggplot2)
library(sdmTMB)
```

@@

# add and scale predictors after filtering for survey(s) of interest
data <- all_depth$data %>%
  filter(depth_max>10) %>% # removes data if sensor stayed too near surface
# dplyr::filter(year > 2003) %>% 
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

@@

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
# # choose the spatial range to model
dat <- data # %>% filter(ssid == 4)
nd <- nd # %>% filter(ssid == 4)

@@

Calculate predicted values
```{r}
m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
# m_temp <- readRDS("../VOCC/data/m_temp_allpost2003.rds")

@@

p <- plot_map(predictions , "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```
```{r}
# choose the spatial range to build raster on
predicted1 <- predictions %>% filter(ssid == 1) 
predicted3 <- predictions %>% filter(ssid == 3) 
predicted4 <- predictions %>% filter(ssid == 4) 
predicted16 <- predictions %>% filter(ssid == 16)

p <- plot_map(predicted1 , "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```

@@

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted16, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-wchg.rds")
```

@@

vector_data <- function(climate_data, 
  ssid = NULL,
  input_cell_size = 2, 
  scale_fac = 2,
  indices = c(1, 1, 1, 2, 2, 3, 3, 3),
  thresholds = c(0.75),
  delta_t = 10,
  time_
  ) {

climate_data <- climate_data %>% filter(ssid == ssid)

if (!is.list(climate_data)) {

@@

} else {
rbrick <- list()
for (i in seq_len(climate_data)) {

rbrick[[i]] <- make_raster_brick(climate_data[[i]], scale_fac = scale_fac)
mnraster_brick <- raster::stackApply(rbrick[[i]], indices = indices, fun = mean)

start_raster <- mnraster_brick[[1]]
end_raster <- mnraster_brick[[3]]
var <- list(start_raster = start_raster, end_raster = end_raster)
}

start_data <- list(var[[i]] = var$start_raster[[i]])
end_data <- list(var[[i]] = var$end_raster[[i]])
}

browser()
start_data #
end_data  #

out <- dist_based_vocc(
  start_data = start_data,
  end_data = end_data,
  x = "x",
  y = "y",
  variable_names = c("index_1"), # what the layer within each element is called
  thresholds = thresholds,
  cell_size =  input_cell_size*scal_fac,
  delta_t = delta_t,
  raster = TRUE
  )

slopedat <-  calcslope(rbrick, time_slices = 10) # vocc::calcslope for comparison
out <- left_join(out, slopedat, by = c("x", "y")) %>% select(-icell)
out$C_per_decade <- out$slope * 10 / out$time_scale
out$km_per_decade <- out$distance * 10 / delta_t

}

vector_data (predictions, 
  ssid = 1,
  input_cell_size = 2, 
  scale_fac = 2,
  indices = c(1, 1, 1, 2, 2, 2, 3, 3, 3),
  thresholds = c(0.75),
  delta_t = 10,
  )
```

>>>>

library(dplyr)
library(ggplot2)
library(sdmTMB)
library(gfranges)
```

@@

# add and scale predictors after filtering for survey(s) of interest
data <- all_depth$data %>%
  filter(depth_max>10) %>% # removes data if sensor stayed too near surface
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

@@

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
# # choose the spatial range to model
dat <- data %>% dplyr::filter(year > 2003) #%>% filter(ssid == 4)
nd <- nd # %>% filter(ssid == 4)

@@

Calculate predicted values
```{r}
m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")

# can also be filtered to include only year with multiple surveys 
# m_temp <- readRDS("../VOCC/data/m_temp_allpost2003.rds")
# nd <- nd %>% dplyr::filter(year > 2003)

@@

p <- plot_map(predictions , "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```
```{r}
# choose the spatial range to build raster on
predicted1 <- predictions %>% filter(ssid == 1) 
predicted3 <- predictions %>% filter(ssid == 3) 
predicted4 <- predictions %>% filter(ssid == 4) 
predicted16 <- predictions %>% filter(ssid == 16) 
predicted1n3 <- predictions %>% filter(ssid != 16) %>% filter(ssid != 4) %>% filter(year > 2004)

p <- plot_map(predicted1n3 , "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```

@@

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted16, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-wchg.rds")

rbrick <- make_raster_brick(predicted1n3, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-hs-qcs.rds")

@@

make_vector_data <- function(climate_data, 
  ssid = NULL,
  start_year = NULL,
  input_cell_size = 2, 
  scale_fac = 2,
  delta_t_total = 10,
  delta_t_step = 2,
  time_slices = 7, 
  indices = c(1, 1, 2, 2, 2, 3, 3),
  thresholds = c(0.75)
  ) {
#browser()
if (!is.null(ssid)) climate_data <- climate_data[climate_data$ssid %in% ssid, ]
if (!is.null(start_year)) climate_data <- climate_data %>% dplyr::filter(year >= start_year)

#if (!identical(time_slices, length(indices)))
  length_indices <- length(indices)
  if(!isTRUE(time_slices==length_indices))
      stop("Must have an indice assigned to each time slice,",
      "therefore length('indices') must equal 'time_slices'.",
      call. = FALSE)

#FIXME: need way of determining how many variables(in separate dataframes) are in dataset  
#if (!is.list(climate_data)) {
# if (length(thresholds) > 1)
#       stop("If multiple climate variables (and corresponding thresholds),", 
#         "data must be in list form.", call. = FALSE)

@@

# } else {
#     if (!identical(length(climate_data), length(thresholds)))
#       stop("Must have a 'thresholds' value for each climate variable", call. = FALSE)
# rbrick <- list()
# for (i in seq_len(climate_data)) {
#   
# rbrick[[i]] <- make_raster_brick(climate_data[[i]], scale_fac = scale_fac)
# mnraster_brick <- raster::stackApply(rbrick[[i]], indices = indices, fun = mean)
# 
# start_raster <- mnraster_brick[[1]]
# end_raster <- mnraster_brick[[3]]
# var <- list(start_raster = start_raster, end_raster = end_raster)
# }

# start_data <- list(var = var$start_raster[[i]])
# end_data <- list(var = var$end_raster[[i]])
#}

start_data #
end_data  #

out <- dist_based_vocc(
  start_data = start_data,
  end_data = end_data,
  x = "x",
  y = "y",
  variable_names = c("index_1"), # what the layer within each element is called
  thresholds = thresholds,
  cell_size =  input_cell_size*scal_fac,
  delta_t = delta_t_total,
  raster = TRUE
  )

# FIXME: need to change function to deal with different time steps within a brick
slopedat <-  calcslope(rbrick, delta_t = delta_t_step) # vocc::calcslope for comparison
out <- left_join(out, slopedat, by = c("x", "y")) %>% select(-icell)
out$C_per_decade <- out$slope * 10
out$km_per_decade <- (out$distance/delta_t)*10 
out
}

make_vector_data (predictions, 
  ssid = c(1, 3),
  start_year = 2005,
  input_cell_size = 2, 
  scale_fac = 2,
  delta_t_total = 10,
  delta_t_step = 2,
  time_slices = 7,
  indices = c(1, 1, 2, 2, 2, 3, 3),
  thresholds = c(0.75)
  )
```
